---
title: "mrubiog2000_OriginalHomeworkCode_03"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(curl)

f <- f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall21/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)

```

QUESTION 1

```{r}
summary(d) #With "summary()" I can see the mean of every quantitative value from the whole population.

attach(d) #Used to be able to call the variables from the data frame.
names(d) #To see the names of the columns (variables) I need to work with.
mean(height)
mean(weight)
mean(zombies_killed)
mean(years_of_education)
mean(age)
#Another way to do it If wanted to see the mean from a specific column (variable).

ssq <- function(x) {sum((x-mean(x))^2)} #Function to calculate the sum of squares.
pop_var <- function(x) {ssq(x)/length(x)} #Function to calculate the variance of the population (note it is divided by "n" instead of by "n-1")
pop_sd <- function(x) {sqrt(pop_var(x))} #Function to calculate the standard deviation of the population.

pop_sd(height)
pop_sd(weight)
pop_sd(zombies_killed)
pop_sd(years_of_education)
pop_sd(age)

#I have calculate the standard deviation of each numeric parameter of the population with the function above. I do not know if there is any function/way to directly apply a function only to numeric values from a data frame, so I did it individually.

```
NK: I had a similar question about applying functions to only numeric values. I don't have the answer, but I ended just selecting the quantitative values from the data frame using select() and then working with the new data frame. However, from a statistical standpoint, I'm not certain I used the correct sd calculations. 

QUESTION 2

```{r}

library(ggplot2)
library(gridExtra) #Package that allows drawing multiple ggplot plots

weight_plot <- ggplot(data = d, aes(x = gender, y = weight)) + geom_boxplot()
height_plot <- ggplot(data = d, aes(x = gender, y = height)) + geom_boxplot()
zombies_plot <- ggplot(data = d, aes(x = gender, y = zombies_killed)) + geom_boxplot()
yearsedu_plot <- ggplot(data = d, aes(x = gender, y = years_of_education)) + geom_boxplot()
age_plot <- ggplot(data = d, aes(x = gender, y = age)) + geom_boxplot()

#Boxplots with gender (x axis) and quantitative variables (y axis)

grid.arrange(weight_plot, height_plot, zombies_plot, yearsedu_plot, age_plot, ncol=3)

#function used to draw multiple plots ("par()" function did not work).

```
NK: for some reason grid.arrange() doesn't work for me, but otherwise we had the same approach here. 

```{r}
#This is an idea I tried first but did not work. Anyways, I think the structure may be useful (selecting only "female" and "male" data from the data frame).

fem <- d[grep("^[Female]", d$gender), ] 
male <- d[grep("^[Male]", d$gender), ] 


#I have selected and separated from the data frame "d" the individuals whose gender is "female" and "male". I tried at first without the "^" but it did not work so I wrote it although I do not know what does it exactly do.
```

QUESTION 3

```{r}

age_height_plot <- ggplot(data = d, aes(x = age, y = height), color = cyl) + geom_point()

age_weight_plot <- ggplot(data = d, aes(x = age, y = weight), color = cyl) + geom_point()

grid.arrange(age_height_plot, age_weight_plot, ncol = 2)

#These variables (age-height, age-weight) might be related in a proportional way (they in/decrease at the same time), but in the age-weight plot there is more dispersion. Anyways, I do not know if there is a way to calculate "R^2" in R

```

QUESTION 4 

```{r}
par(mfrow=c(2,5)) #Used to draw all the plots at the same time (in 2 rows and 5 columns)

hist(d$height, breaks = seq(from = 54, to = 81, by = 1), col = "light green", main = "Height")
hist(d$weight, col = "light green", main = "weight")
hist(d$zombies_killed, col = "red", main = "Zombies killed")
hist(d$years_of_education, col = "light blue", main = "Years of education")
hist(d$age, col = "light blue", main = "Age")

#Histograms of each quantitative variable

qqnorm(d$height, main = "Height Q-Q")
qqline(d$height, col = "red")

qqnorm(d$weight, main = "Weight Q-Q")
qqline(d$weight, col = "red")

qqnorm(d$zombies_killed, main = "Zombies killed Q-Q")
qqline(d$zombies_killed, col = "red")

qqnorm(d$years_of_education, main = "Years of education Q-Q")
qqline(d$years_of_education, col = "red")

qqnorm(d$age, main = "Age Q-Q")
qqline(d$age, col = "red")

#Q-Q plot of each quantitative variable

#I would say that the distributions of "Zombies killed" and "Years of education" are not normal because the data in the Q-Q plot does not match with the line. The shape of their distribution matches with the Poisson distribution (and it seems logic because they are discrete variables that are counted)

```
NK: We came to the same conclusion, although I forgot to comment on which common distribution the data matches. This makes sense though.

QUESTION 5

```{r}
library(dplyr)

set.seed(1) #Used to return the same set of random numbers every time I run the code.
sam1 <- sample_n(d, 30, replace = FALSE) #"Sample_n" function from "dplyr" package allow to randomly select samples from a data frame. I tried to use the "Sample" but id did not work.
sam1

summary(sam1) #To see the mean.
mean(sam1$height) #Another way to calculate the mean directly.
mean(sam1$weight)
mean(sam1$zombies_killed)
mean(sam1$years_of_education)
mean(sam1$age)

sd(sam1$height)
sd(sam1$weight)
sd(sam1$zombies_killed)
sd(sam1$years_of_education)
sd(sam1$age)

#Standard Deviations. I think this time we can use the "sd()" function because we are studying a sample. In QUESTION 1 I think we could not because we were studying the whole population.

se <- function(x) {sd(x)/sqrt(length(x))} #Function to calculate the standard error.
se(sam1$height)
se(sam1$weight)
se(sam1$zombies_killed)
se(sam1$years_of_education)
se(sam1$age)

quantile(sam1$height, c(0.025, 0.975)) 
quantile(sam1$weight, c(0.025, 0.975))
quantile(sam1$age, c(0.025, 0.975)) 

#CI for variables that follow the normal distribution.

IC_Poisson <- function(x) {
  upper = mean(x) + 1.96 * sqrt(mean(x)/length(x))
  lower = mean(x) - 2.58 * sqrt(mean(x)/length(x))
  ci <- c(lower, upper)
  return(ci)
  }

#Function to calculate the IC for a variable that follows Poisson's distribution.
  
  #mean(x) = lambda from Poisson's distribution.
  #This IC can be calculated by multiplying the standard error "sqrt(mean(x)/lenght(x))" by "z*(α/2)" which is 1.96 or 2.58 (I do not understand where these numbers comes from, but when I googled it those were the ones that always appeared).

IC_Poisson(sam1$zombies_killed)
IC_Poisson(sam1$years_of_education)

#CI for variables that follow the Poisson's distribution.

```
NK: I used bootstrapping rather than the Poisson, which gave me a slightly narrower confidence interval. I'm not sure what this says about my method, but our answers are otherwise similar. 

QUESTION 6

```{r}

lol <- NULL
set.seed(1)
for (i in 1:99) {lol[[i]] <- sample_n(d, size = 30, replace = TRUE)}

#for loop to create 99 samples of 30 individuals (zombie survivors).
#Replace = TRUE because if FALSE we do not have enough individuals for 99 samples of 30.

```

```{r}

#I have not been able to calculate the mean of the quantitative variables from the 99 for loop: I do not know how to just do the mean of the variables I want, because the "$" option did not work. I also try to include the mean calculation within the for loop but it did not work either. The book and google did not help either.

#The only think that worked was doing a 99 for loop sampling of each quantitative variable. However, I could not find a way to calculate the mean of each of the 99 subsets.

library(dplyr)

h <- NULL
w <- NULL
z <- NULL
y <- NULL
a <- NULL

set.seed(1)
for (i in 1:99) { h[[i]] <- sample(d$height, size = 30, replace = TRUE) }
for (i in 1:99) { w[[i]] <- sample(d$weight, size = 30, replace = TRUE) }
for (i in 1:99) { z[[i]] <- sample(d$zombies_killed, size = 30, replace = TRUE) }
for (i in 1:99) { y[[i]] <- sample(d$years_of_education, size = 30, replace = TRUE) }
for (i in 1:99) { a[[i]] <- sample(d$age, size = 30, replace = TRUE) }

mean(h[[1]])
mean(w[[1]])
mean(z[[1]])
mean(y[[1]])
mean(a[[1]])

#This only gives the mean from the first subset (out of the 99). I tried the "[[1:99]]" structure but it did not work either.

#I could not answer the remaining questions because I did not obtain the means or the standard variation. I will wait to peer commentary to see if my partners can help me with this (really annoying) problem :)

```
NK: I included the mean calculation in my for loop so the new data frames were just vectors of the 99 means. Then I used mean() and sd() on those data sets.

```{r}
#PROBLEM 1: Calculating the IC for Poisson distribution. I did not understand why I had to use a different method from the Normal distribution. "SOLUTION" The way I discovered in Google worked (I think) but I did not fully understood it.

#PROBLEM 2: It was hard to find a function to draw all the plots at the same time. SOLUTION: I googled it and finally discovered one.

#PROBLEM 3: I was not able to finish Question number 6. I do not know how to apply different functions within the same loop or applying a function to only specific variables to a generated sample from a loop (or just selecting them).

```

